---
title: "NMF"
author: "William Zhang, Eva, Jerry"
date: "2024-09-01"
output:
  pdf_document: default
  html_document: default
---
```{r setup, warning = FALSE, message = FALSE}
# load the packages
library(NMF)
library(tidyverse)
library(gridExtra)
library(readxl)
```

# Procedure
1. Remove hourly observation with missing observation for any chemical
2. Remove background noise level using min values (except for chemicals with minimum value < 2\*LOD and maximum value > 100\*LOD)
3. Zero values are converted to a random value between 0 and 0.5\*LOD
4. Normalize using min and max
5. Compute weight matrix according to Guha's paper, without LOQ

## Reading the data
```{r}
# read the radon data
# Old: 
# hourly_radon <- readRDS("hourly_radon.rds")
# New:
hourly_data <- readRDS("../DataProcessing/Trailer_hourly_merge_20240905.rds")
```

```{r}
# PROCEDURE STEP 1:
hourly_data <- hourly_data %>% rename('co2' = 'co2_ppm')

vocs <- c("ethane", "ethene", "propane", "propene",
                                        "1_3-butadiene", "i-butane", "n-butane",
                                        "acetylene", "cyclopentane", "i-pentane",
                                        "n-pentane", "n-hexane", "isoprene", "n-heptane",
                                        "benzene", "n-octane", "toluene", "ethyl-benzene", 
                                        "m&p-xylene", "o-xylene")


non_vocs <- c('ch4', 'co2', 'co', 'h2s', 'so2', 'nox', 'o3')

# remove row with missing obs for any chemical
hourly_nona <- hourly_data %>% 
  select(c('day', 'time_utc', vocs, non_vocs, 'wdr_deg', 'wsp_ms')) %>% 
  na.omit()

# retrieving the vocs, removing everything else except the vocs
hourly_vocs <- hourly_nona %>% select(any_of(vocs))

# retrieving the non-vocs: co2_ppm, nox, ch4, h2s, so2, o3
# double check this
hourly_non_vocs <- hourly_nona %>% select(all_of(non_vocs)) 

hourly_full_nona <- cbind(hourly_non_vocs, hourly_vocs)

# retrive a vector of yearmonth
hourly_dates <- hourly_nona %>% 
  mutate(yearmonth = substring(day, 0, 7)) %>%
  pull(yearmonth)
``` 

## Data visualisation
```{r fig.height=10}
non_vocs <- c('ch4', 'co2', 'co', 'h2s', 'so2', 'nox', 'o3')
for (compound in non_vocs) {
  assign(paste0(compound, '_boxplot'), 
         ggplot(hourly_non_vocs) +
           geom_boxplot(aes(y = .data[[compound]])) +
           labs(x = compound, y = '') +
           theme_minimal())
}

grid.arrange(ch4_boxplot, co2_boxplot, co_boxplot, nrow = 1)
```

```{r fig.height=10}
grid.arrange(h2s_boxplot, so2_boxplot, nox_boxplot, o3_boxplot, nrow = 1)
```

```{r}
for (compound in vocs) {
  assign(paste0(compound, '_boxplot'), 
         ggplot(hourly_vocs) +
           geom_boxplot(aes(y = .data[[compound]])) +
           labs(x = compound, y = '') +
           theme_minimal())
}
grid.arrange(get(paste0(vocs[1], '_boxplot')), get(paste0(vocs[2], '_boxplot')), 
             get(paste0(vocs[3], '_boxplot')), get(paste0(vocs[4], '_boxplot')), nrow = 1)
```

```{r}
grid.arrange(get(paste0(vocs[5], '_boxplot')), get(paste0(vocs[6], '_boxplot')), 
             get(paste0(vocs[7], '_boxplot')), get(paste0(vocs[8], '_boxplot')), nrow = 1)
```

```{r}
grid.arrange(get(paste0(vocs[9], '_boxplot')), get(paste0(vocs[10], '_boxplot')), 
             get(paste0(vocs[11], '_boxplot')), get(paste0(vocs[12], '_boxplot')), nrow = 1)
```

```{r}
grid.arrange(get(paste0(vocs[13], '_boxplot')), get(paste0(vocs[14], '_boxplot')), 
             get(paste0(vocs[15], '_boxplot')), get(paste0(vocs[16], '_boxplot')), nrow = 1)
```

```{r}
grid.arrange(get(paste0(vocs[17], '_boxplot')), get(paste0(vocs[18], '_boxplot')), 
             get(paste0(vocs[19], '_boxplot')), get(paste0(vocs[20], '_boxplot')), nrow = 1)
```

### Data preprocessing
```{r, message = FALSE}
# Define LOD for each chemical
LOD_non_voc <- c('ch4' = 0.9, 
             'co2' = 0.0433, 
             'co' = 40,
             'h2s' = 0.4, 
             'so2' = 0.4,
             'nox' = 0.05, 
             'o3' = 1)

LOD_voc_monthly <- read_csv('../data/LNM_VOC_LOD_Rounded.csv') %>% select(-1)
# extract the yearmonth from date variables
LOD_voc_monthly <- LOD_voc_monthly %>%
  mutate(yearmonth = strftime(as.POSIXct(start_date, format = '%Y-%m-%d %H:%M:%S', tz = 'UTC'), '%Y-%m')) 

LOD_voc_monthly <- LOD_voc_monthly %>% 
  select(-c(start_date, end_date)) %>%
  select(!any_of(ends_with('half_ldl')))

colnames(LOD_voc_monthly) <- str_replace_all(names(LOD_voc_monthly), '_ldl', '')

LOD_voc_avg <- read_xlsx('../data/LNM_VOC_Uncertainties.xlsx', skip = 1)
LOD_voc_avg <- LOD_voc_avg %>%
  select(1, 4) %>%
  rename('LOD' = 2, 'chemical' = 1) %>%
  head(20)
```

```{r}
# find the min for background-levels
background_levels <- sapply(hourly_full_nona, min)
background_levels
```

```{r}
get_info <- function(column) {
  N <- length(column)
  background <- quantile(column, 0)
  quantile1 <- quantile(column, 0.01)
  quantile99 <- quantile(column, 0.99)
  n_background <- sum(column == background)
  max <- max(column)
  return(c(N, quantile1, quantile99, max, background, n_background))
}

info_table <- hourly_full_nona %>%
  reframe(across(everything(), ~ get_info(.x)))

info_table <- info_table %>%  
  mutate(rownames = c('N', '1st percentile', '99th percentile', 'Max', 'Background', '# Background')) %>%
  pivot_longer(-rownames) %>%   
  pivot_wider(names_from = rownames, values_from = value) 

knitr::kable(info_table)
```

```{r}
# PROCEDURE STEP 2:
#adjustments that were made according to paper
#William: I'm guessing this refers to Gunnar's paper section 2.2 and Guha 3.3

# Check whether chemical has background noise level that needs to be removed
# i.e, NO ADJUSTMENT if minimum value < 2*LOD and maximum value > 100*LOD
adjusting_neg_bg_from_lod <- function(chemical, LOD, background, hourly_data){ 
    # get min and max
    min_value <- min(hourly_data[chemical], na.rm = TRUE)
    max_value <- max(hourly_data[chemical], na.rm = TRUE)
    # if min less than double LOD or max > 100 times LOD
    # adjust to -100 (for entire column???)
    if (min_value < 2 * LOD & max_value > 100 * LOD ){
      return (0)
    }
  return (background)
}
```

```{r}
# Check if background is negligible for non voc
# merge background and LOD
background_lod_non_voc <- tibble(chemical = non_vocs,
                                 LOD = LOD_non_voc,
                                 background = unname(background_levels[non_vocs]))
adjusted_background_non_voc <- background_lod_non_voc %>%
  rowwise() %>%
  mutate(min = min(hourly_full_nona[chemical], na.rm = TRUE),
         LODx2 = 2 * LOD,
         criterion1 = min(hourly_full_nona[chemical], na.rm = TRUE) < 2 * LOD,
         max = max(hourly_full_nona[chemical], na.rm = TRUE),
         LODx100 = 100 * LOD,
         criterion2 = max(hourly_full_nona[chemical], na.rm = TRUE) > 100 * LOD,
         adjusted_background = adjusting_neg_bg_from_lod(chemical, LOD, background, hourly_full_nona))
```

```{r}
# Check if background is negligible for voc
# merge background and LOD
background_lod_voc <- LOD_voc_avg %>%
  left_join(tibble(chemical = setdiff(names(background_levels), non_vocs),
                   background = background_levels[setdiff(names(background_levels), non_vocs)]))
adjusted_background_voc <- background_lod_voc %>%
  rowwise() %>%
  mutate(min = min(hourly_full_nona[chemical], na.rm = TRUE),
         LODx2 = 2 * LOD,
         criterion1 = min(hourly_full_nona[chemical], na.rm = TRUE) < 2 * LOD,
         max = max(hourly_full_nona[chemical], na.rm = TRUE),
         LODx100 = 100 * LOD,
         criterion2 = max(hourly_full_nona[chemical], na.rm = TRUE) > 100 * LOD,
         adjusted_background = adjusting_neg_bg_from_lod(chemical, LOD, background, hourly_full_nona))
```
```{r}
# So now we have the adjusted background concentrations
hourly_nona_bgrm <- hourly_full_nona %>%
  mutate(across(adjusted_background_non_voc$chemical, ~  .x - adjusted_background_non_voc$adjusted_background[adjusted_background_non_voc$chemical== cur_column()]))
hourly_nona_bgrm <- hourly_nona_bgrm %>%
  mutate(across(adjusted_background_voc$chemical, ~  .x - adjusted_background_voc$adjusted_background[adjusted_background_voc$chemical== cur_column()]))
```

```{r}
# look at zero values
colSums(hourly_nona_bgrm == 0)
```

```{r}
# PROCEDURE STEP 3
# replace zero values with random values between 0 and 0.5*LOD
set.seed(123)
replace_zero_with_random <- function(column, name, LOD_df){
  LOD <- LOD_df$LOD[LOD_df$chemical == name]
  column <- if_else(column == 0, round(runif(length(column), 0, 0.5 * LOD), 3), column)
  return (column)
}

hourly_nona_bgrm_zerorepl <- hourly_nona_bgrm %>%
  mutate(across(adjusted_background_non_voc$chemical, ~ replace_zero_with_random(.x, cur_column(), adjusted_background_non_voc)))

hourly_nona_bgrm_zerorepl <- hourly_nona_bgrm_zerorepl %>%
  mutate(across(adjusted_background_voc$chemical, ~ replace_zero_with_random(.x, cur_column(), adjusted_background_voc)))
```


### Normalize the non-vocs
```{r}
#normalizing function
normalize_column <- function(column){
  background <- quantile(column, 0)
  max <- quantile(column, 1) # this could be adjusted
  return ((column - background)/(max - background))
}
```

```{r}
# normalize all
hourly_nona_bgrm_zerorepl_norm <- as_tibble(sapply(as.list(hourly_nona_bgrm_zerorepl), normalize_column)) #normalize the NON_VOC
summary(hourly_nona_bgrm_zerorepl_norm)
```

### Combine and Transpose
```{r}
normalized_matrix <- as.matrix(hourly_nona_bgrm_zerorepl_norm) #important: using the normalized VOCs for this file

# Transpose <- cbind(Normalized_Data, Merged_VOCs)
# rownames(Transpose) <- as.character(Transpose[,1]) # I'm not able to run this line, but it shouldn't affect anything
# transpose_normalized_matrix <- t(as.matrix(normalized_matrix))

number_row<- dim(normalized_matrix)[1] #store number of rows (used for checking)
number_column<- dim(normalized_matrix)[2] #store number of columns
```

## NMF section
```{r}
# compute weight matrix (uncertainties)
# Based on the Guha paper
# next comment is from the other nmf R file
weight_matrix <- matrix(0, nrow = nrow(normalized_matrix), ncol = ncol(normalized_matrix))
LOD_merged <- tibble(chemical = c(adjusted_background_non_voc$chemical, adjusted_background_voc$chemical),
                     LOD = c(adjusted_background_non_voc$LOD, adjusted_background_voc$LOD))

LOD_merged <- tibble(chemical = names(hourly_nona_bgrm_zerorepl_norm)) %>%
  left_join(LOD_merged)

# creating uncertainty Matrix
for (i in 1:number_row) { 
  for (j in 1:number_column) {
    xij <- normalized_matrix[i, j]
    LOD <- LOD_merged$LOD[[j]]
    # Get LOD value for this row 
    if (j == 1) { 
      # based on equation 6, we sqrt ch4 (at column = 1) and times by 1
      weight_matrix[i, j] <- sqrt(xij)
    } else if (j == 2) {
      # 0.25 for co2 
      weight_matrix[i, j] <- 0.25 * sqrt(xij)
    } else if (j == 3) { 
      # 0.5 for CO
      weight_matrix[i, j] <- 0.5 * sqrt(xij)
    } else if (xij <= LOD) {
      weight_matrix[i, j] <- 2 * LOD # equation 5a) in reference paper
    } else {
      weight_matrix[i, j] <- sqrt(((0.1 * xij)**2 + LOD**2))  #equation 5c) in reference paper
    }
  }
}
```

```{r, include = FALSE}
# set a seed for nmf
# set.seed(123)
# #function below used to estimate the optimal rank and will be used in the nmf() function. 
# # takes around 20-30 mins to run
# estimate_rank <- nmfEstimateRank(normalized_matrix, 4:8, method = "ls-nmf", weight = weight_matrix, 30)
# # # changing the range of rank to 2:20 from 4:20
# saveRDS(estimate_rank, 'estimate_rank.rds')

estimate_rank <- readRDS('estimate_rank.rds')
measures <- estimate_rank$measures
fit <- estimate_rank$fit
consensus <- estimate_rank$consensus
```

```{r}
# plots the NMF rank survey
plot(estimate_rank)
```


```{r, include = FALSE}
# fitting the optimal rank based on the above plots
# the choice of the optimal rank needs to be discussed
output <- nmf(normalized_matrix, rank = 4, weight = weight_matrix, method = "ls-nmf")
W <- basis(output)
H <- coef(output)
```

### Source contributions
```{r, include = FALSE}
# Convert H to a data frame for ggplot
H_df <- as.data.frame(H)
# Add a column for component
H_df$Component <- names(as.data.frame(W))

# Reshape data to long format
H_long <- pivot_longer(H_df, cols = -Component, names_to = "Chemical", values_to = "Contribution")

# Define the desired order of chemicals
desired_order <- c(
  # NMHCs - Alkanes
  "ethane", "propane", "i-butane", "n-butane", "i-pentane", "n-pentane", 
  "n-hexane", "cyclopentane", "n-heptane", "n-octane",
  
  # NMHCs - Alkenes
  "ethene", "propene", "1_3-butadiene", 'isoprene',
  
  # NMHCs - Alkynes
  "acetylene",
  
  # NMHCs - Aromatics
  "benzene", "toluene", "ethyl-benzene", "o-xylene", "m&p-xylene",
  
  # Inorganic Gases - CO and CO2
  "co", "co2",
  
  # Nitrogen Oxides (NOx)
  "nox",
  
  # Sulfur Compounds
  "h2s", "so2",
  
  # Ozone (if included)
  "o3",
  
  # Methane
  "ch4"
)

color_pal <- c("#ccd5ae", "#e9edc9", "#fefae0", "#faedcd", "#d4a373", "#df7d5a")

get_component_plot <- function(data, component, title) {
  col <- color_pal[as.numeric(component)]
  component_data <- subset(data, Component == component)
  component_data <- component_data %>%
    mutate(Chemical = factor(Chemical, levels = desired_order))
  
  plot <- ggplot(component_data, aes(x = Chemical, y = Contribution)) +
            geom_bar(stat = "identity", position = "dodge", fill = col) +
            theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
            geom_text(aes(label = sprintf("%.2f", round(Contribution, 2))), color = "blue", size = 3, nudge_y = 0.1) +
            labs(x = "Chemical", y = "Contribution", title = title)+
            theme(
            text = element_text(size = 14), # Base text size for all text elements
            axis.title = element_text(size = 16), # Size of axis titles
            axis.text = element_text(size = 12), # Size of axis text (tick labels)
            plot.title = element_text(size = 18) # Size of the plot title
            )
  return(plot)
}

nmfplt_1_ls <- get_component_plot(H_long, 'V1', 'Component 1 ls-nmf Full')
nmfplt_1_ls
```

```{r, include = FALSE}
nmfplt_2_ls <- get_component_plot(H_long, 'V2', 'Component 2 ls-nmf Full')
nmfplt_2_ls
nmfplt_3_ls <- get_component_plot(H_long, 'V3', 'Component 3 ls-nmf Full')
nmfplt_3_ls
nmfplt_4_ls <- get_component_plot(H_long, 'V4', 'Component 4 ls-nmf Full')
nmfplt_4_ls
```

```{r, include = FALSE, fig.height=15, fig.width=20}
grid.arrange(nmfplt_1_ls, nmfplt_2_ls, nmfplt_3_ls, nmfplt_4_ls, ncol=2)
```

# NMF - SVD seed
```{r}
# Try Eva's approach
components <- 4:10
errors <- numeric(length(components) - 4)

# Loop over the number of components
# for (n in components) {
#   nmf_result <- nmf(normalized_matrix, rank = n, method = "KL", seed='nndsvd')
#   reconstruction <- basis(nmf_result) %*% coef(nmf_result)
#   error <- norm(normalized_matrix - reconstruction, type = "F")
#   errors[n-3] <- error
#   print(paste0('Completed ', n - 3, ' out of 7'))
# }
# 
# saveRDS(errors, 'errors_norm.rds')

errors <- readRDS('errors_norm.rds')
```


```{r, eval=TRUE, echo = F}
# Plot the errors
df <- data.frame(components, errors)
ggplot(df, aes(x = components, y = errors)) +
  geom_line() +
  geom_point() +
  theme_minimal() +
  geom_text(aes(label = sprintf("%.3f", errors)), color = "blue", size = 3, vjust = -0.5) +
  labs(title = "NMF Reconstruction Error vs. Number of Components",
       x = "Number of Components", y = "Reconstruction Error")
```


```{r, include = FALSE}
rankerror4 <-  df[1,2]
rankerror4
```

```{r, include = FALSE}
# Apply NMF
nmf_result <- nmf(normalized_matrix, rank = 4, method = "KL", seed='nndsvd')

# Examine the basis matrix (W)
basis_matrix <- basis(nmf_result)
# print(Inyodata_basis_matrix)

# Examine the coefficient matrix (H)
coef_matrix <- coef(nmf_result)

# Optionally, visualize the results
par(mfrow = c(1, 2))
# Basis matrix
image(basis_matrix, main = "Basis Matrix (W)")
# Coefficient matrix
image(coef_matrix, main = "Coefficient Matrix (H)")
```


```{r, include=FALSE}
# Convert H to a data frame for ggplot
H_df <- as.data.frame(coef_matrix)
# Add a column for chemicals
H_df$Component <- rownames(H_df)

# Reshape data to long format
H_long <- pivot_longer(H_df, cols = -Component, names_to = "Chemical", values_to = "Contribution")

# Plot
nmfplt_1_svd <- get_component_plot(H_long, '1', 'Component 1 nndsvd Full')
nmfplt_1_svd
```

```{r, include=FALSE}
nmfplt_2_svd <- get_component_plot(H_long, '2', 'Component 2 nndsvd Full')
nmfplt_3_svd <- get_component_plot(H_long, '3', 'Component 3 nndsvd Full')
nmfplt_4_svd <- get_component_plot(H_long, '4', 'Component 4 nndsvd Full')
```

```{r, include = FALSE, fig.height=15, fig.width=16}
grid.arrange(nmfplt_1_svd, nmfplt_2_svd, nmfplt_3_svd, nmfplt_4_svd, ncol=2)
```
\newpage

# Method comparisons
```{r, include = FALSE, fig.height=8, fig.width=16}
grid.arrange(nmfplt_1_ls, nmfplt_3_svd, ncol = 2)
```

```{r, include = FALSE, fig.height=8, fig.width=16}
grid.arrange(nmfplt_2_ls, nmfplt_2_svd, ncol = 2)
```

```{r, include = FALSE, fig.height=8, fig.width=16}
grid.arrange(nmfplt_3_ls, nmfplt_1_svd, ncol = 2)
```

```{r, include = FALSE, fig.height=8, fig.width=16}
grid.arrange(nmfplt_4_ls, nmfplt_4_svd, ncol = 2)
```

# Remove Ozone

## 4 Components
```{r}
normalized_matrix_less_o3 <- normalized_matrix[ ,setdiff(colnames(normalized_matrix), "o3")]
# Store this for other use
write_csv(as_tibble(normalized_matrix_less_o3), 'normalized_matrix_less_o3.csv')

nmf_result_4c_less_o3 <- nmf(normalized_matrix_less_o3, rank = 4, method = "KL", seed='nndsvd')

basis_matrix_4c_less_o3 <- basis(nmf_result_4c_less_o3)
coef_matrix_4c_less_o3 <- coef(nmf_result_4c_less_o3)

par(mfrow = c(1, 2))
image(basis_matrix_4c_less_o3, main = "Basis Matrix (W)")
image(coef_matrix_4c_less_o3, main = "Coefficient Matrix (H)")
```

```{r, eval=TRUE, echo = T}
# Convert H to a data frame for ggplot
H_df_4c_less_o3 <- as.data.frame(coef_matrix_4c_less_o3)
# Add a column for chemicals
H_df_4c_less_o3$Component <- rownames(H_df_4c_less_o3)

# Reshape data to long format
H_long_4c_less_o3 <- pivot_longer(H_df_4c_less_o3, cols = -Component, names_to = "Chemical", values_to = "Contribution")

# Plot
nmfplt_1_svd_4c_less_o3 <- get_component_plot(H_long_4c_less_o3, 
                                           '1', 'Component 1 nndsvd Less ozone, 4C')
nmfplt_2_svd_4c_less_o3 <- get_component_plot(H_long_4c_less_o3, 
                                           '2', 'Component 2 nndsvd Less ozone, 4C')
nmfplt_3_svd_4c_less_o3 <- get_component_plot(H_long_4c_less_o3, 
                                           '3', 'Component 3 nndsvd Less ozone, 4C')
nmfplt_4_svd_4c_less_o3 <- get_component_plot(H_long_4c_less_o3, 
                                           '4', 'Component 4 nndsvd Less ozone, 4C')
```

```{r, eval=TRUE, echo = FALSE, fig.height=15, fig.width=16}
grid.arrange(nmfplt_1_svd_4c_less_o3, nmfplt_2_svd_4c_less_o3, nmfplt_3_svd_4c_less_o3, nmfplt_4_svd_4c_less_o3, ncol=2)
```

## 5 Components
```{r, message = FALSE}
nmf_result_5c_less_o3 <- nmf(normalized_matrix_less_o3, rank = 5, method = "KL", seed='nndsvd')

basis_matrix_5c_less_o3 <- basis(nmf_result_5c_less_o3)
coef_matrix_5c_less_o3 <- coef(nmf_result_5c_less_o3)

par(mfrow = c(1, 2))
image(basis_matrix_5c_less_o3, main = "Basis Matrix (W)")
image(coef_matrix_5c_less_o3, main = "Coefficient Matrix (H)")
```

```{r, eval=TRUE, echo = T}
# Convert H to a data frame for ggplot
H_df_5c_less_o3 <- as.data.frame(coef_matrix_5c_less_o3)
# Add a column for chemicals
H_df_5c_less_o3$Component <- rownames(H_df_5c_less_o3)

# Reshape data to long format
H_long_5c_less_o3 <- pivot_longer(H_df_5c_less_o3, cols = -Component, names_to = "Chemical", values_to = "Contribution")

# Plot
nmfplt_1_svd_5c_less_o3 <- get_component_plot(H_long_5c_less_o3, 
                                           '1', '1) Traffic emissions factor')
nmfplt_2_svd_5c_less_o3 <- get_component_plot(H_long_5c_less_o3, 
                                           '2', '2) Component 2 nndsvd Less ozone, 5c')
nmfplt_3_svd_5c_less_o3 <- get_component_plot(H_long_5c_less_o3, 
                                           '3', '3) Oil & Gas emssions factor')
nmfplt_4_svd_5c_less_o3 <- get_component_plot(H_long_5c_less_o3, 
                                           '4', '4) Flaring factor II (other directions)')
nmfplt_5_svd_5c_less_o3 <- get_component_plot(H_long_5c_less_o3, 
                                           '5', '5) Flaring Factor I (SW pad\'s flare)')
```

```{r, eval=TRUE, echo = FALSE, fig.height=15, fig.width=16}
grid.arrange(nmfplt_1_svd_5c_less_o3, nmfplt_2_svd_5c_less_o3, nmfplt_3_svd_5c_less_o3, nmfplt_4_svd_5c_less_o3, nmfplt_5_svd_5c_less_o3, ncol=2)
```

### Wind plots
```{r}
hourly_wind_nona <- hourly_nona %>%
  select(wdr_deg, wsp_ms)

data_to_plot <- tibble(
  component1 = basis(nmf_result_5c_less_o3)[,1],
  component2 = basis(nmf_result_5c_less_o3)[,2],
  component3 = basis(nmf_result_5c_less_o3)[,3],
  component4 = basis(nmf_result_5c_less_o3)[,4],
  component5 = basis(nmf_result_5c_less_o3)[,5],
  wd = round(hourly_wind_nona$wdr_deg, -1)
)

```

```{r, eval=TRUE, echo = FALSE}
boxplot(component1 ~ wd, data = data_to_plot,
        outline = FALSE, col = color_pal[1],
        main = "NMF factor expression vs Wind Direction (Component 1)",
        xlab = "Wind direction (°)",
        ylab = "Factor expression")
```

```{r, eval=TRUE, echo = FALSE}
boxplot(component2 ~ wd, data = data_to_plot,
        outline = FALSE, col = color_pal[2],
        main = "NMF factor expression vs Wind Direction (Component 2)",
        xlab = "Wind direction (°)",
        ylab = "Factor expression")
```

```{r, eval=TRUE, echo = FALSE}
boxplot(component3 ~ wd, data = data_to_plot,
        outline = FALSE, col = color_pal[3],
        main = "NMF factor expression vs Wind Direction (Component 3)",
        xlab = "Wind direction (°)",
        ylab = "Factor expression")
```

```{r, eval=TRUE, echo = FALSE}
boxplot(component4 ~ wd, data = data_to_plot,
        outline = FALSE, col = color_pal[4],
        main = "NMF factor expression vs Wind Direction (Component 4)",
        xlab = "Wind direction (°)",
        ylab = "Factor expression")
```

```{r, eval=TRUE, echo = FALSE}
boxplot(component5 ~ wd, data = data_to_plot,
        outline = FALSE, col = color_pal[5],
        main = "NMF factor expression vs Wind Direction (Component 5)",
        xlab = "Wind direction (°)",
        ylab = "Factor expression")

```
## Factor analysis
```{r}
# First look at how well this approximates
fitted_5c_less_o3 <- fitted(nmf_result_5c_less_o3)
sum(abs(normalized_matrix_less_o3-fitted_5c_less_o3))
```
```{r, eval=TRUE, echo = T}
# NMF factorizes V = WH
# Store Basis matrix (W) and Coef Matrix (H)
saveRDS(basis_matrix_5c_less_o3, 'result_rfiles/nmf_norm_5c_less_o3_basis.rds')
saveRDS(coef_matrix_5c_less_o3, 'result_rfiles/nmf_norm_5c_less_o3_coef.rds')

# Merge basis matrix into hourly observations
basis_matrix_5c_less_o3 <- as_tibble(basis_matrix_5c_less_o3) %>%
  setNames(c('Factor1', 'Factor2', 'Factor3', 'Factor4', 'Factor5'))

normalized_hourly_data_5c_less_o3 <- hourly_nona[,c('day', 'time_utc')] %>%
  cbind(normalized_matrix_less_o3) %>%
  cbind(basis_matrix_5c_less_o3) %>%
  right_join(hourly_data %>% select(-'day'), join_by(time_utc), suffix = c('_norm', ''))

# saveRDS(normalized_full_data_5c_less_o3, 'result_rfiles/normalized_hourly_data_5c_less_o3.rds')
```

```{r}
# Also compute a daily dataset
normalized_daily_data_5c_less_o3 <- normalized_hourly_data_5c_less_o3 %>%
  group_by(day) %>%
  summarise(across(where(is.numeric), ~ mean(.x, na.rm = T)))

# saveRDS(normalized_daily_data_5c_less_o3, 'result_rfiles/normalized_daily_data_5c_less_o3.rds')
```


```{r}
# Check if relationship between # flares and flare factor (4 & 5)

# Linear model
flare_factor <- lm(n_flare_100 ~ Factor4 + Factor5, data = normalized_daily_data_5c_less_o3)
summary(flare_factor)

flare_factor_weighted <- lm(weighted.count ~ Factor4 + Factor5, data = normalized_daily_data_5c_less_o3)
summary(flare_factor_weighted)

# Poisson model
flare_factor_pois <- glm(n_flare_100 ~ Factor4 + Factor5, family = 'poisson', data = normalized_daily_data_5c_less_o3)
summary(flare_factor_pois)
```

```{r}
# Check relationship between avg flare distance and flare factor (4 & 5)
# Linear model
flare_factor_dist <- lm(distToLovi ~ Factor4 + Factor5, data = normalized_daily_data_5c_less_o3)
summary(flare_factor_dist)
```



## 6 Components
```{r, include = FALSE}
nmf_result_6c_less_o3 <- nmf(normalized_matrix_less_o3, rank = 6, method = "KL", seed='nndsvd')

basis_matrix_6c_less_o3 <- basis(nmf_result_6c_less_o3)
coef_matrix_6c_less_o3 <- coef(nmf_result_6c_less_o3)

par(mfrow = c(1, 2))
image(basis_matrix_6c_less_o3, main = "Basis Matrix (W)")
image(coef_matrix_6c_less_o3, main = "Coefficient Matrix (H)")
```

```{r, eval=TRUE, echo = T, include=FALSE}
# Convert H to a data frame for ggplot
H_df_6c_less_o3 <- as.data.frame(coef_matrix_6c_less_o3)
# Add a column for chemicals
H_df_6c_less_o3$Component <- rownames(H_df_6c_less_o3)

# Reshape data to long format
H_long_6c_less_o3 <- pivot_longer(H_df_6c_less_o3, cols = -Component, names_to = "Chemical", values_to = "Contribution")

# Plot
nmfplt_1_svd_6c_less_o3 <- get_component_plot(H_long_6c_less_o3, 
                                           '1', 'Component 1 nndsvd Less ozone, 6c')
nmfplt_2_svd_6c_less_o3 <- get_component_plot(H_long_6c_less_o3, 
                                           '2', 'Component 2 nndsvd Less ozone, 6c')
nmfplt_3_svd_6c_less_o3 <- get_component_plot(H_long_6c_less_o3, 
                                           '3', 'Component 3 nndsvd Less ozone, 6c')
nmfplt_4_svd_6c_less_o3 <- get_component_plot(H_long_6c_less_o3, 
                                           '4', 'Component 4 nndsvd Less ozone, 6c')
nmfplt_5_svd_6c_less_o3 <- get_component_plot(H_long_6c_less_o3, 
                                           '5', 'Component 5 nndsvd Less ozone, 6c')
nmfplt_6_svd_6c_less_o3 <- get_component_plot(H_long_6c_less_o3, 
                                           '6', 'Component 6 nndsvd Less ozone, 6c')
```

```{r, eval=TRUE, echo = FALSE, include = FALSE, fig.height=15, fig.width=16}
grid.arrange(nmfplt_1_svd_6c_less_o3, nmfplt_2_svd_6c_less_o3, nmfplt_3_svd_6c_less_o3, nmfplt_4_svd_6c_less_o3, nmfplt_5_svd_6c_less_o3, nmfplt_6_svd_6c_less_o3, ncol=2)
```

### Wind plots
```{r, include = FALSE}
data_to_plot <- tibble(
  component1 = basis(nmf_result_6c_less_o3)[,1],
  component2 = basis(nmf_result_6c_less_o3)[,2],
  component3 = basis(nmf_result_6c_less_o3)[,3],
  component4 = basis(nmf_result_6c_less_o3)[,4],
  component5 = basis(nmf_result_6c_less_o3)[,5],
  component6 = basis(nmf_result_6c_less_o3)[,6],
  wd = round(hourly_wind_nona$wdr_deg, -1)
)
```

```{r, include = FALSE}
boxplot(component1 ~ wd, data = data_to_plot,
        outline = FALSE, col = "#ccd5ae",
        main = "NMF factor expression vs Wind Direction (Component 1)",
        xlab = "Wind direction (°)",
        ylab = "Factor expression")
```

```{r, include = FALSE}
boxplot(component2 ~ wd, data = data_to_plot,
        outline = FALSE, col = "#e9edc9",
        main = "NMF factor expression vs Wind Direction (Component 2)",
        xlab = "Wind direction (°)",
        ylab = "Factor expression")
```

```{r, include = FALSE}
boxplot(component3 ~ wd, data = data_to_plot,
        outline = FALSE, col = "#fefae0",
        main = "NMF factor expression vs Wind Direction (Component 3)",
        xlab = "Wind direction (°)",
        ylab = "Factor expression")
```

```{r, include = FALSE}
boxplot(component4 ~ wd, data = data_to_plot,
        outline = FALSE, col = "#faedcd",
        main = "NMF factor expression vs Wind Direction (Component 4)",
        xlab = "Wind direction (°)",
        ylab = "Factor expression")
```

```{r, include = FALSE}
boxplot(component5 ~ wd, data = data_to_plot,
        outline = FALSE, col = "#d4a373",
        main = "NMF factor expression vs Wind Direction (Component 5)",
        xlab = "Wind direction (°)",
        ylab = "Factor expression")

```

```{r, include = FALSE}
boxplot(component6 ~ wd, data = data_to_plot,
        outline = FALSE, col = "#f0a373",
        main = "NMF factor expression vs Wind Direction (Component 6)",
        xlab = "Wind direction (°)",
        ylab = "Factor expression")

```

# Remove Ozone + chemicals with more than 500+ background values
## 4 Components
```{r, include = FALSE}
normalized_matrix_less_o3_lotbg <- normalized_matrix[ ,setdiff(colnames(normalized_matrix), c('o3', 'h2s', 'so2', '1_3-butadiene', 'isoprene', 'ethyl-benzene', 'm&p-xylene', 'o-xylene'))]

nmf_result_4c_less_o3_lotbg <- nmf(normalized_matrix_less_o3_lotbg, rank = 4, method = "KL", seed='nndsvd')

basis_matrix_4c_less_o3_lotbg <- basis(nmf_result_4c_less_o3_lotbg)
coef_matrix_4c_less_o3_lotbg <- coef(nmf_result_4c_less_o3_lotbg)

par(mfrow = c(1, 2))
image(basis_matrix_4c_less_o3_lotbg, main = "Basis Matrix (W)")
image(coef_matrix_4c_less_o3_lotbg, main = "Coefficient Matrix (H)")
```

```{r, eval=TRUE, echo = T, include=FALSE}
# Convert H to a data frame for ggplot
H_df_4c_less_o3_lotbg <- as.data.frame(coef_matrix_4c_less_o3_lotbg)
# Add a column for chemicals
H_df_4c_less_o3_lotbg$Component <- rownames(H_df_4c_less_o3_lotbg)

# Reshape data to long format
H_long_4c_less_o3_lotbg <- pivot_longer(H_df_4c_less_o3_lotbg, cols = -Component, names_to = "Chemical", values_to = "Contribution")

# Plot
nmfplt_1_svd_4c_less_o3_lotbg <- get_component_plot(H_long_4c_less_o3_lotbg, 
                                           '1', 'Component 1 nndsvd Less ozone, lot bg, 4C')
nmfplt_2_svd_4c_less_o3_lotbg <- get_component_plot(H_long_4c_less_o3_lotbg, 
                                           '2', 'Component 2 nndsvd Less ozone, lot bg, 4C')
nmfplt_3_svd_4c_less_o3_lotbg <- get_component_plot(H_long_4c_less_o3_lotbg, 
                                           '3', 'Component 3 nndsvd Less ozone, lot bg, 4C')
nmfplt_4_svd_4c_less_o3_lotbg <- get_component_plot(H_long_4c_less_o3_lotbg, 
                                           '4', 'Component 4 nndsvd Less ozone, lot bg, 4C')
```

```{r, eval=TRUE, echo = FALSE, include = FALSE, fig.height=15, fig.width=16}
grid.arrange(nmfplt_1_svd_4c_less_o3_lotbg, nmfplt_2_svd_4c_less_o3_lotbg, nmfplt_3_svd_4c_less_o3_lotbg, nmfplt_4_svd_4c_less_o3_lotbg, ncol=2)
```

## 5 Components
```{r, include = FALSE}
nmf_result_5c_less_o3_lotbg <- nmf(normalized_matrix_less_o3_lotbg, rank = 5, method = "KL", seed='nndsvd')

basis_matrix_5c_less_o3_lotbg <- basis(nmf_result_5c_less_o3_lotbg)
coef_matrix_5c_less_o3_lotbg <- coef(nmf_result_5c_less_o3_lotbg)

par(mfrow = c(1, 2))
image(basis_matrix_5c_less_o3_lotbg, main = "Basis Matrix (W)")
image(coef_matrix_5c_less_o3_lotbg, main = "Coefficient Matrix (H)")
```

```{r, eval=TRUE, echo = T, include=FALSE}
# Convert H to a data frame for ggplot
H_df_5c_less_o3_lotbg <- as.data.frame(coef_matrix_5c_less_o3_lotbg)
# Add a column for chemicals
H_df_5c_less_o3_lotbg$Component <- rownames(H_df_5c_less_o3_lotbg)

# Reshape data to long format
H_long_5c_less_o3_lotbg <- pivot_longer(H_df_5c_less_o3_lotbg, cols = -Component, names_to = "Chemical", values_to = "Contribution")

# Plot
nmfplt_1_svd_5c_less_o3_lotbg <- get_component_plot(H_long_5c_less_o3_lotbg, 
                                           '1', 'Component 1 nndsvd Less ozone, lot bg, 5c')
nmfplt_2_svd_5c_less_o3_lotbg <- get_component_plot(H_long_5c_less_o3_lotbg, 
                                           '2', 'Component 2 nndsvd Less ozone, lot bg, 5c')
nmfplt_3_svd_5c_less_o3_lotbg <- get_component_plot(H_long_5c_less_o3_lotbg, 
                                           '3', 'Component 3 nndsvd Less ozone, lot bg, 5c')
nmfplt_4_svd_5c_less_o3_lotbg <- get_component_plot(H_long_5c_less_o3_lotbg, 
                                           '4', 'Component 4 nndsvd Less ozone, lot bg, 5c')
nmfplt_5_svd_5c_less_o3_lotbg <- get_component_plot(H_long_5c_less_o3_lotbg, 
                                           '5', 'Component 5 nndsvd Less ozone, lot bg, 5c')
```

```{r, eval=TRUE, echo = FALSE, include = FALSE, fig.height=15, fig.width=16}
grid.arrange(nmfplt_1_svd_5c_less_o3_lotbg, nmfplt_2_svd_5c_less_o3_lotbg, nmfplt_3_svd_5c_less_o3_lotbg, nmfplt_4_svd_5c_less_o3_lotbg, nmfplt_5_svd_5c_less_o3_lotbg, ncol=2)
```
Compare to 4 components

```{r, eval=TRUE, echo = FALSE, include = FALSE, fig.height=15, fig.width=16}
grid.arrange(nmfplt_1_svd_4c_less_o3_lotbg, nmfplt_2_svd_4c_less_o3_lotbg, nmfplt_3_svd_4c_less_o3_lotbg, nmfplt_4_svd_4c_less_o3_lotbg, ncol=2)
```