---
title: "NMF Final (Diurnal))"
author: "William Zhang, Eva, Jerry, Meredith"
date: "2025-03-21"
output:
  html_document: default
  pdf_document: default
---
```{r setup, warning = FALSE, message = FALSE}
# load the packages
library(NMF)
library(tidyverse)
library(grid)
library(gridExtra)
library(readxl)
library(circular)
library(lwgeom)
library(units)
```

# Procedure
1. Remove hourly observation with missing observation for any chemical
2. Remove background noise level using min values (except for chemicals with minimum value < 2\*LOD and maximum value > 100\*LOD)
3. Zero values are converted to a random value between 0 and 0.5\*LOD
4. Normalize using min and max
5. Remove Ozone (wouldn't affect # of obs.)

## Reading the data
```{r}
hourly_data <- readRDS("../DataProcessing/Trailer_hourly_merge_20240905.rds")
```

```{r}
# PROCEDURE STEP 1:
hourly_data_diurnal <- hourly_data %>% rename('co2' = 'co2_ppm') %>%
  mutate(datetime_mountain = with_tz(as.POSIXct(time_utc, tz = 'UTC', 
                                                format = "%Y-%m-%d %H:%M:%OS"), 
                                     tzone = "America/Denver")) %>%
  filter(date(datetime_mountain) >= '2023-05-01') %>%
  filter(hour(datetime_mountain) >= 11 & hour(datetime_mountain) <= 17)

vocs <- c("ethane", "ethene", "propane", "propene",
                                        "1_3-butadiene", "i-butane", "n-butane",
                                        "acetylene", "cyclopentane", "i-pentane",
                                        "n-pentane", "n-hexane", "isoprene", "n-heptane",
                                        "benzene", "n-octane", "toluene", "ethyl-benzene", 
                                        "m&p-xylene", "o-xylene")


non_vocs <- c('ch4', 'co2', 'co', 'h2s', 'so2', 'nox', 'o3')

# remove row with missing obs for any chemical
hourly_nona_diurnal <- hourly_data_diurnal %>% 
  select(any_of(c('day', 'time_utc', vocs, non_vocs, 'wdr_deg', 'wsp_ms'))) %>% 
  na.omit()

# retrieving the vocs, removing everything else except the vocs
hourly_vocs_diurnal <- hourly_nona_diurnal %>% select(any_of(vocs))

# retrieving the non-vocs: co2_ppm, nox, ch4, h2s, so2, o3
# double check this
hourly_non_vocs_diurnal <- hourly_nona_diurnal %>% select(any_of(non_vocs)) 

hourly_full_nona_diurnal <- cbind(hourly_non_vocs_diurnal, hourly_vocs_diurnal)

# retrive a vector of yearmonth
hourly_dates_diurnal <- hourly_nona_diurnal %>% 
  mutate(yearmonth = substring(day, 0, 7)) %>%
  pull(yearmonth)
``` 

## Data visualisation

- Boxplots of the hourly concentrations non-voc

```{r, echo = FALSE, out.width="70%"}
non_vocs <- c('ch4', 'co2', 'co', 'h2s', 'so2', 'nox', 'o3')
for (compound in non_vocs) {
  assign(paste0(compound, '_boxplot'), 
         ggplot(hourly_non_vocs_diurnal) +
           geom_boxplot(aes(x = .data[[compound]])) +
           labs(y = compound, x = '') +
           theme_minimal())
}

grid.arrange(ch4_boxplot, co2_boxplot, co_boxplot, nrow = 3)
```

- Boxplots sulfur compounds, NOx, ozone

```{r, echo = FALSE, out.width="70%"}
grid.arrange(h2s_boxplot, so2_boxplot, nox_boxplot, o3_boxplot, nrow = 4)
```

- Boxplots VOCs

```{r, echo = FALSE, out.width="70%"}
for (compound in vocs) {
  assign(paste0(compound, '_boxplot'), 
         ggplot(hourly_vocs_diurnal) +
           geom_boxplot(aes(x = .data[[compound]])) +
           labs(y = compound, x = '') +
           theme_minimal())
}
grid.arrange(get(paste0(vocs[1], '_boxplot')), get(paste0(vocs[2], '_boxplot')), 
             get(paste0(vocs[3], '_boxplot')), get(paste0(vocs[4], '_boxplot')), nrow = 4)
```

```{r, echo = FALSE, out.width="70%"}
grid.arrange(get(paste0(vocs[5], '_boxplot')), get(paste0(vocs[6], '_boxplot')), 
             get(paste0(vocs[7], '_boxplot')), get(paste0(vocs[8], '_boxplot')), nrow = 4)
```

```{r, echo = FALSE, out.width="70%"}
grid.arrange(get(paste0(vocs[9], '_boxplot')), get(paste0(vocs[10], '_boxplot')), 
             get(paste0(vocs[11], '_boxplot')), get(paste0(vocs[12], '_boxplot')), nrow = 4)
```

```{r, echo = FALSE, out.width="70%"}
grid.arrange(get(paste0(vocs[13], '_boxplot')), get(paste0(vocs[14], '_boxplot')), 
             get(paste0(vocs[15], '_boxplot')), get(paste0(vocs[16], '_boxplot')), nrow = 4)
```

```{r, echo = FALSE, out.width="70%"}
grid.arrange(get(paste0(vocs[17], '_boxplot')), get(paste0(vocs[18], '_boxplot')), 
             get(paste0(vocs[19], '_boxplot')), get(paste0(vocs[20], '_boxplot')), nrow = 4)
```

### Data pre-processing

#### Step 1: limits of detection
- STEP 1: Limits of detection

```{r, message = FALSE}
# Define LOD for each chemical
LOD_non_voc <- c('ch4' = 0.9, 
             'co2' = 0.0433, 
             'co' = 20,
             'h2s' = 0.4, 
             'so2' = 0.4,
             'nox' = 0.05, 
             'o3' = 1)

# LOD_voc_monthly <- read_csv('../data/LNM_VOC_LOD_Rounded.csv') %>% select(-1)
# 
# # extract the yearmonth from date variables
# LOD_voc_monthly <- LOD_voc_monthly %>%
#   mutate(yearmonth = strftime(as.POSIXct(start_date, format = '%Y-%m-%d %H:%M:%S', 
#                                             tz = 'UTC'), '%Y-%m')) 
# 
# LOD_voc_monthly <- LOD_voc_monthly %>% 
#   select(-c(start_date, end_date)) %>%
#   select(!any_of(ends_with('half_ldl')))
# 
# colnames(LOD_voc_monthly) <- str_replace_all(names(LOD_voc_monthly), '_ldl', '')

LOD_voc_avg <- read_xlsx('../data/LNM_VOC_Uncertainties.xlsx', skip = 1)
LOD_voc_avg <- LOD_voc_avg %>%
  select(1, 4) %>%
  rename('LOD' = 2, 'chemical' = 1) %>%
  head(20)
```

#### Step 2: Background correction
- STEP 2: Background correction 

```{r echo=FALSE}
# find the min for background-levels
background_levels <- sapply(hourly_full_nona_diurnal, min)
background_levels
```
- Summary statistics of backgrounds and extremes

```{r}
get_info <- function(column) {
  N <- length(column)
  background <- quantile(column, 0)
  quantile1 <- quantile(column, 0.01)
  quantile99 <- quantile(column, 0.99)
  n_background <- sum(column == background)
  max <- max(column)
  return(c(N, quantile1, quantile99, max, background, n_background))
}

info_table <- hourly_full_nona_diurnal %>%
  reframe(across(everything(), ~ get_info(.x)))

info_table <- info_table %>%  
  mutate(rownames = c('N', '1st percentile', '99th percentile', 'Max', 
                      'Background', '# Background')) %>%
  pivot_longer(-rownames) %>%   
  pivot_wider(names_from = rownames, values_from = value) 

knitr::kable(info_table)
```

- STEP 2 processing continued: background correction
- adjustments that were made according to paper: Gunnar's paper section 2.2 and Guha 3.3
- Check whether chemical has background noise level that needs to be removed
- NO ADJUSTMENT if minimum value < 2xLOD and maximum value > 100xLOD

```{r}
adjusting_neg_bg_from_lod <- function(chemical, LOD, background, hourly_data_diurnal){ 
    # get min and max
    min_value <- min(hourly_data_diurnal[chemical], na.rm = TRUE)
    max_value <- max(hourly_data_diurnal[chemical], na.rm = TRUE)
    # if min less than double LOD or max > 100 times LOD
    # adjust to -100 (for entire column???)
    if (min_value < 2 * LOD & max_value > 100 * LOD ){
      return (0)
    }
  return (background)
}
```

- Check if background is negligible for non voc
- merge background and LOD

```{r}
background_lod_non_voc_diurnal <- tibble(chemical = non_vocs,
                                 LOD = LOD_non_voc,
                                 background = unname(background_levels[non_vocs]))
adjusted_background_non_voc_diurnal <- background_lod_non_voc_diurnal %>%
  rowwise() %>%
  mutate(min = min(hourly_full_nona_diurnal[chemical], na.rm = TRUE),
         LODx2 = 2 * LOD,
         criterion1 = min(hourly_full_nona_diurnal[chemical], na.rm = TRUE) < 2 * LOD,
         max = max(hourly_full_nona_diurnal[chemical], na.rm = TRUE),
         LODx100 = 100 * LOD,
         criterion2 = max(hourly_full_nona_diurnal[chemical], na.rm = TRUE) > 100 * LOD,
         adjusted_background = adjusting_neg_bg_from_lod(chemical, LOD, background, 
                                                         hourly_full_nona_diurnal))
```


- Check if background is negligible for voc
- merge background and LOD


```{r message=FALSE, warning=FALSE}
background_lod_voc_diurnal <- LOD_voc_avg %>%
  left_join(tibble(chemical = setdiff(names(background_levels), non_vocs),
                   background = background_levels[setdiff(names(background_levels), 
                                                          non_vocs)]))
adjusted_background_voc_diurnal <- background_lod_voc_diurnal %>%
  rowwise() %>%
  mutate(min = min(hourly_full_nona_diurnal[chemical], na.rm = TRUE),
         LODx2 = 2 * LOD,
         criterion1 = min(hourly_full_nona_diurnal[chemical], na.rm = TRUE) < 2 * LOD,
         max = max(hourly_full_nona_diurnal[chemical], na.rm = TRUE),
         LODx100 = 100 * LOD,
         criterion2 = max(hourly_full_nona_diurnal[chemical], na.rm = TRUE) > 100 * LOD,
         adjusted_background = adjusting_neg_bg_from_lod(chemical, LOD, background, 
                                                         hourly_full_nona_diurnal))
```
- create dataset with background removed

```{r}
# So now we have the adjusted background concentrations
hourly_nona_bgrm_diurnal <- hourly_full_nona_diurnal %>%
  mutate(across(adjusted_background_non_voc_diurnal$chemical, 
                ~  .x - adjusted_background_non_voc_diurnal$adjusted_background[
                    adjusted_background_non_voc_diurnal$chemical == cur_column()]))
hourly_nona_bgrm_diurnal <- hourly_nona_bgrm_diurnal %>%
  mutate(across(adjusted_background_voc_diurnal$chemical, 
                ~  .x - adjusted_background_voc_diurnal$adjusted_background[
                    adjusted_background_voc_diurnal$chemical == cur_column()]))
```

- check number of 0 values per compound

```{r}
# look at zero values
colSums(hourly_nona_bgrm_diurnal == 0)
```

#### Step 3: Replace zero with random value between 0 and 0.5*LOD
- STEP 3: replace zero values with random values between 0 and 0.5xLOD

```{r}
set.seed(123)
replace_zero_with_random <- function(column, name, LOD_df){
  LOD <- LOD_df$LOD[LOD_df$chemical == name]
  column <- if_else(column == 0, round(runif(length(column), 0, 0.5 * LOD), 3), column)
  return (column)
}

hourly_nona_bgrm_zerorepl_diurnal <- hourly_nona_bgrm_diurnal %>%
  mutate(across(adjusted_background_non_voc_diurnal$chemical,
                ~ replace_zero_with_random(.x, cur_column(), adjusted_background_non_voc_diurnal)))

hourly_nona_bgrm_zerorepl_diurnal <- hourly_nona_bgrm_zerorepl_diurnal %>%
  mutate(across(adjusted_background_voc_diurnal$chemical,
                ~ replace_zero_with_random(.x, cur_column(), adjusted_background_voc_diurnal)))
```

#### Step 4: Normalize
- STEP 4: Normalize the measurements

```{r}
#normalizing function
normalize_column <- function(column){
  background <- quantile(column, 0)
  max <- quantile(column, 1) # this could be adjusted
  return ((column - background)/(max - background))
}
```

```{r}
# normalize all
hourly_nona_bgrm_zerorepl_norm_diurnal <- as_tibble(sapply(as.list(hourly_nona_bgrm_zerorepl_diurnal),
                                                   normalize_column))
#normalize the NON_VOC
summary(hourly_nona_bgrm_zerorepl_norm_diurnal)
```

- FINAL step: create matrix of processed and normalized concentrations for NMF

```{r}
normalized_matrix <- as.matrix(hourly_nona_bgrm_zerorepl_norm_diurnal)
#important: using the normalized VOCs for this file
```

# NMF section
## Preprocess
### Global variables
```{r}
components <- 4:10
```

### Remove Ozone
```{r}
normalized_matrix_less_o3 <- normalized_matrix[ ,setdiff(colnames(normalized_matrix), "o3")]
```

### Compute error matrix
```{r}
# compute uncertainty matrix (inverse of weight?)
# Based on the Guha paper

uncertainty_matrix <- matrix(0, nrow = nrow(normalized_matrix_less_o3), 
                        ncol = ncol(normalized_matrix_less_o3))
LOD_merged <- tibble(chemical = c(adjusted_background_non_voc_diurnal$chemical, 
                                  adjusted_background_voc_diurnal$chemical),
                     LOD = c(adjusted_background_non_voc_diurnal$LOD, 
                             adjusted_background_voc_diurnal$LOD))

LOD_merged <- tibble(chemical = names(hourly_nona_bgrm_zerorepl_norm_diurnal)) %>%
  left_join(LOD_merged) %>%
  filter(chemical %in% colnames(normalized_matrix_less_o3))

# creating uncertainty Matrix
for (i in 1:dim(uncertainty_matrix)[1]) { 
  for (j in 1:dim(uncertainty_matrix)[2]) {
    chemical <- colnames(normalized_matrix_less_o3)[j]
    xij <- normalized_matrix_less_o3[i, j]
    LOD <- LOD_merged$LOD[LOD_merged$chemical == chemical]
    # Get LOD value for this row 
    # Based on Guha Eq6, Eq5a, EQ5c
    # if (j == 1) { 
    #   # based on equation 6, we sqrt ch4 (at column = 1) and times by 1
    #   uncertainty_matrix[i, j] <- sqrt(xij)
    # } else if (j == 2) {
    #   # 0.25 for co2 
    #   uncertainty_matrix[i, j] <- 0.25 * sqrt(xij)
    # } else if (j == 3) { 
    #   # 0.5 for CO
    #   uncertainty_matrix[i, j] <- 0.5 * sqrt(xij)
    # } else if (xij <= LOD) {
    #   uncertainty_matrix[i, j] <- 2 * LOD # equation 5a) in reference paper
    # } else {
    #   uncertainty_matrix[i, j] <- sqrt(((0.1 * xij)**2 + LOD**2))  #equation 5c) in reference paper
    # }
    
    # Based on Guha Eq5a, EQ5c
    if (xij <= LOD) {
      uncertainty_matrix[i, j] <- 2 * LOD # equation 5a) in reference paper
    } else {
      uncertainty_matrix[i, j] <- sqrt(((0.1 * xij)**2 + LOD**2))  #equation 5c) in reference paper
    }
  }
}
```

```{r}
# Convert zero uncertainties to the next smallest uncertainty of the corresponding compound
uncertainty_matrix[uncertainty_matrix==0]<-apply(uncertainty_matrix, 2, function(x) sort(x)[2])
# THIS NEEDS TO BE CHECKED IF WE WANT TO TAKE RECIPROCAL FOR EACH ELEMENT
# CURRENT RESULTS IS WHEN WEIGHT = UNCERTAINTY
# NOT POSSIBLE TO DO SIMPLY TAKE RECIPROCAL SINCE THERE'RE 0 UNCERTAINTIES
weight_matrix <- 1/uncertainty_matrix 

summary(weight_matrix)
```

### Helper functions for plots

```{r, include = FALSE}
# Define the desired order of chemicals
desired_order <- c(
  # NMHCs - Alkanes
  "ethane", "propane", "i-butane", "n-butane", "i-pentane", "n-pentane", 
  "n-hexane", "cyclopentane", "n-heptane", "n-octane",
  
  # NMHCs - Alkenes
  "ethene", "propene", "1_3-butadiene", 'isoprene',
  
  # NMHCs - Alkynes
  "acetylene",
  
  # NMHCs - Aromatics
  "benzene", "toluene", "ethyl-benzene", "o-xylene", "m&p-xylene",
  
  # Inorganic Gases - CO and CO2
  "co", "co2",
  
  # Nitrogen Oxides (NOx)
  "nox",
  
  # Sulfur Compounds
  "h2s", "so2",
  
  # Ozone (if included)
  "o3",
  
  # Methane
  "ch4"
)

color_pal<-c("#00AFBB", "#E7B800", "#FC4E07","#0072B2","#8B4513")

get_component_plot <- function(data, component, title) {
  col <- color_pal[as.numeric(component)]
  component_data <- subset(data, Component == component)
  component_data <- component_data %>%
    mutate(Chemical = factor(Chemical, levels = desired_order))
  
  plot <- ggplot(component_data, aes(x = Chemical, y = Contribution)) +
            geom_bar(stat = "identity", position = "dodge", fill = col) +
            theme_minimal()+
            theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
            geom_text(aes(label = sprintf("%.2f", round(Contribution, 2))), 
                      color = "blue", size = 3, vjust = -0.5) +
            labs(x = "Chemical", y = "Contribution", title = title)+
            theme(
            text = element_text(size = 14), # Base text size for all text elements
            axis.title = element_text(size = 16), # Size of axis titles
            axis.text = element_text(size = 12), # Size of axis text (tick labels)
            plot.title = element_text(size = 18) # Size of the plot title
            )
  return(plot)
}
```

```{r}
get_fingerprint_plot <- function(H, factor_names = c('Factor 1', 'Factor 2',
                                                     'Factor 3', 'Factor 4',
                                                     'Factor 5')){
  custom_colors <- setNames(color_pal,
                            factor_names)

  # Convert to proportions
  contrib_prop <- apply(H[,1:(length(H)-1)], MARGIN = 2, 
                        FUN = function(x) {x/sum(x)})
  
  contrib_prop <- contrib_prop %>%
    as_tibble() %>%
    mutate(Component = factor_names) %>%
    mutate(Component = factor(Component, levels = factor_names)) %>%
    pivot_longer(cols = -Component,
                 names_to = "Chemical",
                 values_to = "Contribution_prop") %>%
    mutate(Chemical = factor(Chemical, levels = desired_order))
    
  return(contrib_prop %>%
    ggplot(aes(fill = Component, y = Contribution_prop, x = Chemical)) +
    geom_bar(position = "fill", stat = "identity") +
    scale_fill_manual(values = custom_colors) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    labs(x = "Chemical", y = "Contribution Proportion") +
    theme(
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      panel.background = element_blank()
    ))
}
```

```{r}
hourly_wind_nona_diurnal <- hourly_nona_diurnal %>%
  select(wdr_deg, wsp_ms)

get_wind_plot_data <- function(W, factor_names = c('Factor 1', 'Factor 2',
                                                     'Factor 3', 'Factor 4',
                                                     'Factor 5')){
  data_to_plot <- tibble(
    component1 = W[, 1],
    component2 = W[, 2],
    component3 = W[, 3],
    component4 = W[, 4],
    component5 = W[, 5],
    wd = round(hourly_wind_nona_diurnal$wdr_deg, -1)
  )
  
  data_long <- data_to_plot %>%
  pivot_longer(cols = starts_with("component"), names_to = "Factor", values_to = "Expression")
  
  data_long <- data_long %>%
    mutate(wd = factor(wd, levels = sort(unique(wd))))  
    
  data_long
}

get_wind_plots <- function(W, y_axis_upper = rep(10, 5),
                               factor_names = c('Factor 1', 'Factor 2',
                                                     'Factor 3', 'Factor 4',
                                                     'Factor 5')){
  data_long <- get_wind_plot_data(W, factor_names)
  
  # Select every second wind direction for labeling
  every_second_label <- levels(data_long$wd)[seq(1, length(levels(data_long$wd)), by = 2)]
  
  factor_labels <- setNames(paste(c('Factor 1', 'Factor 2','Factor 3', 'Factor 4',
                                    'Factor 5'), ' - ', factor_names), paste0('component', 1:5))
    
  y_axis_limits <- list(
    "component1" = c(0, y_axis_upper[1]),
    "component2" = c(0, y_axis_upper[2]),
    "component3" = c(0, y_axis_upper[3]),
    "component4" = c(0, y_axis_upper[4]),
    "component5" = c(0, y_axis_upper[5])
  )


  plots <- lapply(1:5, function(i) {
    factor_name <- paste0("component", i)
    
    ggplot(data_long %>% filter(Factor == factor_name), 
           aes(x = wd, y = Expression, fill = as.factor(wd))) +
      geom_boxplot(outliers=F, size=0.3) +  
      scale_fill_manual(values = rep(color_pal[i], length(unique(data_long$wd)))) +  
      scale_x_discrete(breaks = every_second_label) +  
      coord_cartesian(ylim = y_axis_limits[[factor_name]]) + 
      scale_y_continuous(
        limits = c(0, NA),
        breaks = seq(0, y_axis_limits[[factor_name]][2], length.out = 5) ,
        expand=expansion(mult=c(0))
      ) +
      labs(title = factor_labels[factor_name], 
           x = "Wind Direction (°)", 
           y = "Factor Expression") +
      theme_minimal() +
      theme(
        legend.position = "none",
        plot.title = element_text(size = 6),  # Smaller title text
        axis.title = element_text(size = 6),  # Smaller axis labels
        axis.text = element_text(size = 6),  # Smaller x and y tick labels
        axis.text.x = element_text(angle = 45, hjust = 1)
      )
  })
  
  return(plots)
}
```

## LS-NMF + Random seed
```{r}
# for each rank, run 30 models and find best one
# start_time_lsnmf_rand <- Sys.time()
# 
# lsnmf_random_less_o3 <- nmf(
#   normalized_matrix_less_o3,
#   components,
#   method = "ls-nmf",
#   weight = weight_matrix,
#   30,
#   seed = 123456
# )
# 
# end_time_lsnmf_rand <- Sys.time()
# end_time_lsnmf_rand-start_time_lsnmf_rand
# 5.1 minutes to run the above

# saveRDS(lsnmf_random_less_o3,
#         'lsnmf_random_less_o3_diurnal.rds')
lsnmf_random_less_o3 <- readRDS('lsnmf_random_less_o3_diurnal.rds')
```

```{r}
# plots the NMF rank survey
plot(lsnmf_random_less_o3)
```

### Look at 5 factors:
```{r}
output <- lsnmf_random_less_o3$fit$`5`
W <- basis(output)
H <- coef(output)
```

#### Source Contribution plots
* Source Contribution plots

```{r}
# Convert H to a data frame for ggplot
H_df_5c_less_o3 <- as.data.frame(H)
# Add a column for chemicals
H_df_5c_less_o3$Component <- rownames(H_df_5c_less_o3)

# Reshape data to long format
H_long_5c_less_o3 <- pivot_longer(H_df_5c_less_o3, cols = -Component, 
                                  names_to = "Chemical", values_to = "Contribution")

# Plot
nmfplt_1_lsnmf_random_less_o3_5c <- get_component_plot(H_long_5c_less_o3, 
                                           '1', '1)')
nmfplt_2_lsnmf_random_less_o3_5c <- get_component_plot(H_long_5c_less_o3, 
                                           '2', '2)')
nmfplt_3_lsnmf_random_less_o3_5c <- get_component_plot(H_long_5c_less_o3, 
                                           '3', '3)')
nmfplt_4_lsnmf_random_less_o3_5c <- get_component_plot(H_long_5c_less_o3, 
                                           '4', '4)')
nmfplt_5_lsnmf_random_less_o3_5c <- get_component_plot(H_long_5c_less_o3, 
                                           '5', '5)')
```


```{r, eval=TRUE, echo = FALSE, fig.height=15, fig.width=16}
grid.arrange(nmfplt_1_lsnmf_random_less_o3_5c, nmfplt_2_lsnmf_random_less_o3_5c, 
             nmfplt_3_lsnmf_random_less_o3_5c, nmfplt_4_lsnmf_random_less_o3_5c, 
             nmfplt_5_lsnmf_random_less_o3_5c, 
             ncol=2, top = textGrob("LS-NMF Random Seed Diurnal",gp=gpar(fontsize=20,font=3)))
#ggsave("factors.png", p)
```

#### Fingerprint plot
* Fingerprint plots

```{r}
fingerprint <- get_fingerprint_plot(H_df_5c_less_o3, c(
      'Component 1',
      'Component 2',
      'Component 3',
      'Component 4',
      'Component 5'
    ))

fingerprint
```

#### Wind plot
* Wind plots

```{r}
wind_plot <- get_wind_plots(W, y_axis_upper = c(3.75, 4.5, 3, 1.75, 2))

grid.arrange(grobs = wind_plot, ncol = 3)
```

## LS-NMF + nndsvd seed
```{r, warning = FALSE}
# Run nmf with 4:10 components and nndsvd seed
# start_time_lsnmf_nndsvd <- Sys.time()
# 
# lsnmf_nndsvd_less_o3 <- nmf(
#   normalized_matrix_less_o3,
#   rank = components,
#   nrun = 1, # since using nndsvd
#   method = "ls-nmf",
#   weight = weight_matrix,
#   seed = 'nndsvd'
# )
# 
# end_time_lsnmf_nndsvd <- Sys.time()
# end_time_lsnmf_nndsvd-start_time_lsnmf_nndsvd
# 20 seconds to run the above
#
# saveRDS(lsnmf_nndsvd_less_o3,
#         'lsnmf_nndsvd_less_o3_diurnal.rds')

lsnmf_nndsvd_less_o3 <- readRDS('lsnmf_nndsvd_less_o3_diurnal.rds')
```

```{r}
# plots the NMF rank survey
plot(lsnmf_nndsvd_less_o3)
```

### Look at 5 factors:
```{r}
output <- lsnmf_nndsvd_less_o3$fit$`5`
W <- basis(output)
H <- coef(output)
```

#### Source Contribution plots
* Source Contribution plots

```{r}
# Convert H to a data frame for ggplot
H_df_5c_less_o3 <- as.data.frame(H)
# Add a column for chemicals
H_df_5c_less_o3$Component <- rownames(H_df_5c_less_o3)

# Reshape data to long format
H_long_5c_less_o3 <- pivot_longer(H_df_5c_less_o3, cols = -Component, 
                                  names_to = "Chemical", values_to = "Contribution")

# Plot
nmfplt_1_lsnmf_nndsvd_less_o3_5c <- get_component_plot(H_long_5c_less_o3, 
                                           '1', '1)')
nmfplt_2_lsnmf_nndsvd_less_o3_5c <- get_component_plot(H_long_5c_less_o3, 
                                           '2', '2)')
nmfplt_3_lsnmf_nndsvd_less_o3_5c <- get_component_plot(H_long_5c_less_o3, 
                                           '3', '3)')
nmfplt_4_lsnmf_nndsvd_less_o3_5c <- get_component_plot(H_long_5c_less_o3, 
                                           '4', '4)')
nmfplt_5_lsnmf_nndsvd_less_o3_5c <- get_component_plot(H_long_5c_less_o3, 
                                           '5', '5)')
```


```{r, eval=TRUE, echo = FALSE, fig.height=15, fig.width=16}
grid.arrange(nmfplt_1_lsnmf_nndsvd_less_o3_5c, nmfplt_2_lsnmf_nndsvd_less_o3_5c, 
             nmfplt_3_lsnmf_nndsvd_less_o3_5c, nmfplt_4_lsnmf_nndsvd_less_o3_5c, 
             nmfplt_5_lsnmf_nndsvd_less_o3_5c, 
             ncol=2, top = textGrob("LS-NMF nndsvd Seed Diurnal",gp=gpar(fontsize=20,font=3)))
#ggsave("factors.png", p)
```

#### Fingerprint plot
* Fingerprint plots

```{r}
fingerprint <- get_fingerprint_plot(H_df_5c_less_o3, c(
      'Component 1',
      'Component 2',
      'Component 3',
      'Component 4',
      'Component 5'
    ))
fingerprint
#ggsave("fingerprint.png", c)
```

#### Wind plot
* Wind plots

```{r}
wind_plot <- get_wind_plots(W, y_axis_upper = c(0.5, 0.2, 0.4, 0.2, 0.25))
grid.arrange(grobs = wind_plot, ncol = 3)
```

## Comparing random seed vs nndsvd for ls-nmf
Residuals is defined as sum(((X - fitted(object)) * weight)^2)/2

```{r}
tibble(component = components,
       random_residual = lsnmf_random_less_o3$measures$residuals,
       nndsvd_residual = lsnmf_nndsvd_less_o3$measures$residuals) %>%
  ggplot() +
  geom_line(aes(x = component, y =random_residual, color = 'Random seed')) +
  geom_point(aes(x = component, y =random_residual, color = 'Random seed')) +
  geom_line(aes(x = component, y = nndsvd_residual, color = 'nndsvd')) +
  geom_point(aes(x = component, y = nndsvd_residual, color = 'nndsvd')) +
  scale_colour_manual("", 
                    breaks = c("Random seed", "nndsvd"),
                    values = c("tomato1", "dodgerblue")) +
  labs(x = 'Rank', y = 'Residual', 'WRSS') +
  theme_bw()
```

## KL + Random seed

```{r, warning = FALSE}
# start_time_kl_random <- Sys.time()
# 
# kl_random_less_o3 <- nmf(
#   normalized_matrix_less_o3,
#   rank = components,
#   nrun = 30,
#   method = "KL",
#   seed = 123456
# )
# 
# end_time_kl_random <- Sys.time()
# end_time_kl_random-start_time_kl_random
# 3.8 minutes to run the above
# 
# saveRDS(kl_random_less_o3, 'kl_random_less_o3_diurnal.rds')

kl_random_less_o3 <- readRDS('kl_random_less_o3_diurnal.rds')
```

```{r}
# plots the NMF rank survey
plot(kl_random_less_o3)
```

### Look at 5 factors:

```{r}
output <- kl_random_less_o3$fit$`5`
W <- basis(output)
H <- coef(output)
```

#### Source Contribution plots
* Source Contribution plots

```{r}
# Convert H to a data frame for ggplot
H_df_5c_less_o3 <- as.data.frame(H)
# Add a column for chemicals
H_df_5c_less_o3$Component <- rownames(H_df_5c_less_o3)

# Reshape data to long format
H_long_5c_less_o3 <- pivot_longer(H_df_5c_less_o3, cols = -Component, 
                                  names_to = "Chemical", values_to = "Contribution")

# Plot
nmfplt_1_kl_random_less_o3_5c <- get_component_plot(H_long_5c_less_o3, 
                                           '1', '1)')
nmfplt_2_kl_random_less_o3_5c <- get_component_plot(H_long_5c_less_o3, 
                                           '2', '2)')
nmfplt_3_kl_random_less_o3_5c <- get_component_plot(H_long_5c_less_o3, 
                                           '3', '3)')
nmfplt_4_kl_random_less_o3_5c <- get_component_plot(H_long_5c_less_o3, 
                                           '4', '4)')
nmfplt_5_kl_random_less_o3_5c <- get_component_plot(H_long_5c_less_o3, 
                                           '5', '5)')
```


```{r, eval=TRUE, echo = FALSE, fig.height=15, fig.width=16}
grid.arrange(nmfplt_1_kl_random_less_o3_5c, nmfplt_2_kl_random_less_o3_5c, 
             nmfplt_3_kl_random_less_o3_5c, nmfplt_4_kl_random_less_o3_5c, 
             nmfplt_5_kl_random_less_o3_5c, 
             ncol=2, top = textGrob("KL Random Seed Diurnal",gp=gpar(fontsize=20,font=3)))
#ggsave("factors.png", p)
```

#### Fingerprint plot
* Fingerprint plots

```{r}
fingerprint <- get_fingerprint_plot(H_df_5c_less_o3, c(
      'Component 1',
      'Component 2',
      'Component 3',
      'Component 4',
      'Component 5'
    ))
fingerprint
#ggsave("fingerprint.png", c)
```

#### Wind plot
* Wind plots

```{r}
wind_plot <- get_wind_plots(W, y_axis_upper = c(1.25, 1.5, 3, 4, 2))
grid.arrange(grobs = wind_plot, ncol = 3)
```

## KL + nndsvd
```{r, warning = FALSE}
# start_time_kl_nndsvd <- Sys.time()
# 
# kl_nndsvd_less_o3 <- nmf(
#   normalized_matrix_less_o3,
#   rank = components,
#   nrun = 1,
#   method = "KL",
#   seed = 'nndsvd'
# )
# 
# end_time_kl_nndsvd <- Sys.time()
# end_time_kl_nndsvd-start_time_kl_nndsvd
# 14.5 seconds to run the above
# 
# saveRDS(kl_nndsvd_less_o3, 'kl_nndsvd_less_o3_diurnal.rds')

kl_nndsvd_less_o3 <- readRDS('kl_nndsvd_less_o3_diurnal.rds')
```

```{r}
# plots the NMF rank survey
plot(kl_nndsvd_less_o3)
```

### Look at 5 factors:

```{r}
output <- kl_nndsvd_less_o3$fit$`5`
W <- basis(output)
H <- coef(output)
```

#### Source Contribution plots
* Source Contribution plots

```{r}
# Convert H to a data frame for ggplot
H_df_5c_less_o3 <- as.data.frame(H)
# Add a column for chemicals
H_df_5c_less_o3$Component <- rownames(H_df_5c_less_o3)

# Reshape data to long format
H_long_5c_less_o3 <- pivot_longer(H_df_5c_less_o3, cols = -Component, 
                                  names_to = "Chemical", values_to = "Contribution")

# Plot
nmfplt_1_kl_nndsvd_less_o3_5c <- get_component_plot(H_long_5c_less_o3, 
                                           '1', '1)')
nmfplt_2_kl_nndsvd_less_o3_5c <- get_component_plot(H_long_5c_less_o3, 
                                           '2', '2)')
nmfplt_3_kl_nndsvd_less_o3_5c <- get_component_plot(H_long_5c_less_o3, 
                                           '3', '3)')
nmfplt_4_kl_nndsvd_less_o3_5c <- get_component_plot(H_long_5c_less_o3, 
                                           '4', '4)')
nmfplt_5_kl_nndsvd_less_o3_5c <- get_component_plot(H_long_5c_less_o3, 
                                           '5', '5)')
```


```{r, eval=TRUE, echo = FALSE, fig.height=15, fig.width=16}
grid.arrange(nmfplt_1_kl_nndsvd_less_o3_5c, nmfplt_2_kl_nndsvd_less_o3_5c, 
             nmfplt_3_kl_nndsvd_less_o3_5c, nmfplt_4_kl_nndsvd_less_o3_5c, 
             nmfplt_5_kl_nndsvd_less_o3_5c, 
             ncol=2, top = textGrob("KL nndsvd Seed Diurnal",gp=gpar(fontsize=20,font=3)))
```

#### Fingerprint plot
* Fingerprint plots

```{r}
fingerprint <- get_fingerprint_plot(H_df_5c_less_o3, c(
      'Component 1',
      'Component 2',
      'Component 3',
      'Component 4',
      'Component 5'
    ))
fingerprint
#ggsave("fingerprint.png", c)
```

#### Wind plot
* Wind plots

```{r}
wind_plot <- get_wind_plots(W, y_axis_upper = c(0.5, 0.3, 0.225, 0.225, 0.25))
grid.arrange(grobs = wind_plot, ncol = 3)
```

```{r, eval=FALSE, echo = FALSE}
boxplot(component1 ~ wd, data = data_to_plot,
        outline = FALSE, col = color_pal[1],
        main = "Factor 1 - Traffic",
        xlab = "Wind direction (°)",
        ylab = "Factor expression")

boxplot(component2 ~ wd, data = data_to_plot,
        outline = FALSE, col = color_pal[2],
        main = "Factor 2 - Combustion",
        xlab = "Wind direction (°)",
        ylab = "Factor expression")

boxplot(component3 ~ wd, data = data_to_plot,
        outline = FALSE, col = color_pal[3],
        main = "Factor 3 - Oil & Gas Emissions",
        xlab = "Wind direction (°)",
        ylab = "Factor expression")

boxplot(component4 ~ wd, data = data_to_plot,
        outline = FALSE, col = color_pal[4],
        main = "Factor 4 - Flaring",
        xlab = "Wind direction (°)",
        ylab = "Factor expression")

boxplot(component5 ~ wd, data = data_to_plot,
        outline = FALSE, col = color_pal[5],
        main = "Factor 5 - Flaring Sour Gas",
        xlab = "Wind direction (°)",
        ylab = "Factor expression")
```


## Comparing all four methods
Using RSS, WRSS, and KL
```{r}
get_residual <- function(component, seed, method, objective) {
  fitted <- fitted(get(paste(method, seed, 'less_o3', sep = '_'))$fit[[component-3]])
  if (objective == 'wrss') {
    return(sum(((normalized_matrix_less_o3 - fitted) * weight_matrix)^2)/2)
  } else if (objective == 'kl') {
    log_term <- normalized_matrix_less_o3/fitted
    log_term[log_term<.Machine$double.eps] <- .Machine$double.eps
    return(sum(normalized_matrix_less_o3 * log(log_term) + fitted - normalized_matrix_less_o3))
  } else if (objective == 'rss') {
    return(norm(normalized_matrix_less_o3 - fitted, type = 'F')^2)
  }
}

df_plot <- expand_grid(
  component = components,
  seed = c('random', 'nndsvd'),
  method = c('lsnmf', 'kl'),
  objective = c('rss', 'wrss', 'kl')
) %>%
  rowwise() %>%
  mutate(residual = get_residual(component, seed, method, objective)) %>%
  ungroup() %>%
  mutate(model = paste0(method, seed))

RSS_plot <- df_plot %>%
  filter(objective=='rss') %>%
  ggplot() +
  geom_line(aes(x = component, y = residual, group = model, color = model)) +
  geom_point(aes(x = component, y = residual, group = model, color = model)) +
  scale_colour_manual("", 
                    breaks = c("lsnmfrandom", "lsnmfnndsvd",
                               "klrandom", "klnndsvd"),
                    values = c("tomato1", "dodgerblue",
                               "springgreen4", "darkorange")) +
  labs(x = 'Rank', y = 'Residual', title = 'RSS') +
  theme_bw() +
  theme(legend.position="none")

WRSS_plot <- df_plot %>%
  filter(objective=='wrss') %>%
  ggplot() +
  geom_line(aes(x = component, y = residual, group = model, color = model)) +
  geom_point(aes(x = component, y = residual, group = model, color = model)) +
  scale_colour_manual("", 
                  breaks = c("lsnmfrandom", "lsnmfnndsvd",
                             "klrandom", "klnndsvd"),
                  values = c("tomato1", "dodgerblue",
                             "springgreen4", "darkorange")) +
  labs(x = 'Rank', y = '', title = 'wrss') +
  theme_bw() +
  theme(legend.position="none")

KL_plot <- df_plot %>%
  filter(objective=='kl') %>%
  ggplot() +
  geom_line(aes(x = component, y = residual, group = model, color = model)) +
  geom_point(aes(x = component, y = residual, group = model, color = model)) +
  scale_colour_manual("", 
                    breaks = c("lsnmfrandom", "lsnmfnndsvd",
                               "klrandom", "klnndsvd"),
                    values = c("tomato1", "dodgerblue",
                               "springgreen4", "darkorange")) +
  labs(x = 'Rank', y = '', title = 'KL') +
  theme_bw()
grid.arrange(RSS_plot, WRSS_plot, KL_plot, ncol=2)
```


## NMF with 5 source factors without ozone

- remove ozone
- use KL divergence loss with svd seed
- Extract W (basis) and H (coefs) matrices 
- Calculate variance explained in all 5 factors
- Calculate variance explained by each factor

```{r, message = FALSE}
nmf_result_5c_less_o3 <- lsnmf_nndsvd_less_o3$fit$`5`

basis_matrix_5c_less_o3 <- basis(nmf_result_5c_less_o3) #W
coef_matrix_5c_less_o3 <- coef(nmf_result_5c_less_o3) #H

# get variance explained by the factors (WRSS)
reconstruct<-fitted(nmf_result_5c_less_o3)

# Traditional way of computing WRSS = sum of weight * (X - Xhat)^2
wtss_trad <- sum(weight_matrix * (normalized_matrix_less_o3 - mean(normalized_matrix_less_o3))^2)
wrss_trad <- sum(weight_matrix * (normalized_matrix_less_o3 - reconstruct)^2)
variance_explained_trad <- 1 - (wrss_trad / wtss_trad)
cat('Variance explained (Traditional WRSS): ', round(variance_explained_trad*100, 2), '%', sep = '')

# How the NMF package computes WRSS = sum of 0.5 * (weight * (X - Xhat))^2
wtss_pkg <- sum((weight_matrix * (normalized_matrix_less_o3 - mean(normalized_matrix_less_o3)))^2/2)
wrss_pkg <- sum((weight_matrix * (normalized_matrix_less_o3 - reconstruct))^2/2)
variance_explained_pkg <- 1 - (wrss_pkg / wtss_pkg)
cat('Variance explained (Package WRSS): ', round(variance_explained_pkg*100, 2), '%', sep = '')


# get variance explained by each factor separately
# Compute variance explained by each factor
# Initialize variance explained tracker
variance_explained_trad_factors <- numeric(5)
variance_explained_pkg_factors <- numeric(5)

# Incrementally add factors and calculate variance explained
reconstruction <- matrix(0, nrow = nrow(basis_matrix_5c_less_o3), ncol = ncol(coef_matrix_5c_less_o3))

for (i in 1:5) {
  # Add the i-th factor to the reconstruction
  reconstruction <- reconstruction + (basis_matrix_5c_less_o3[, i, drop=FALSE] %*% coef_matrix_5c_less_o3[i, , drop=FALSE])
  
  # Compute Residual Sum of Squares (RSS)
  wrss_temp_trad <- sum(weight_matrix * (normalized_matrix_less_o3 - reconstruction)^2)
  wrss_temp_pkg <- sum((weight_matrix * (normalized_matrix_less_o3 - reconstruction))^2/2)
  
  # Compute Variance Explained by adding this factor
  variance_explained_trad_factors[i] <- 1 - (wrss_temp_trad / wtss_trad)
  variance_explained_pkg_factors[i] <- 1 - (wrss_temp_pkg / wtss_pkg)
}

# Plot variance explained by each factor cumulatively
tibble(factor = c(1:length(variance_explained_trad_factors),
                  1:length(variance_explained_pkg_factors)),
       var_exp = c(variance_explained_trad_factors, variance_explained_pkg_factors),
       method = c(rep('Traditional', length(variance_explained_trad_factors)),
                  rep('Package', length(variance_explained_pkg_factors)))) %>%
  ggplot() +
  geom_line(aes(x = factor, y = var_exp, color = method)) +
  geom_point(aes(x = factor, y = var_exp, color = method)) +
  labs(x = 'Number of factors', y = 'Variance explained (WRSS)') +
  theme_bw()
```

```{r, eval=TRUE, echo = T}
# Convert H to a data frame for ggplot
H_df_5c_less_o3 <- as.data.frame(coef_matrix_5c_less_o3)
# Add a column for chemicals
H_df_5c_less_o3$Component <- rownames(H_df_5c_less_o3)

# Reshape data to long format
H_long_5c_less_o3 <- pivot_longer(H_df_5c_less_o3, cols = -Component, 
                                  names_to = "Chemical", values_to = "Contribution")

# Plot
nmfplt_1_svd_5c_less_o3 <- get_component_plot(H_long_5c_less_o3, 
                                           '1', '1) Traffic emissions factor')
nmfplt_2_svd_5c_less_o3 <- get_component_plot(H_long_5c_less_o3, 
                                           '2', '2) Combustion from engines, turbines, compressors')
nmfplt_3_svd_5c_less_o3 <- get_component_plot(H_long_5c_less_o3, 
                                           '3', '3) Oil & gas fugitive and venting emssions')
nmfplt_4_svd_5c_less_o3 <- get_component_plot(H_long_5c_less_o3, 
                                           '4', '4) Flaring, incomplete combustion')
nmfplt_5_svd_5c_less_o3 <- get_component_plot(H_long_5c_less_o3, 
                                           '5', '5) Flaring, sour gas')
```

```{r, eval=TRUE, echo = FALSE, fig.height=15, fig.width=16}
grid.arrange(nmfplt_1_svd_5c_less_o3, nmfplt_2_svd_5c_less_o3, 
             nmfplt_3_svd_5c_less_o3, nmfplt_4_svd_5c_less_o3, nmfplt_5_svd_5c_less_o3, 
             ncol=2)
#ggsave("factors.png", p)
```


## Factor analysis

- merge in factors 1-5 to dataset (hourly)
```{r}
# First look at how well this approximates
fitted_5c_less_o3 <- fitted(nmf_result_5c_less_o3)
sum(abs(normalized_matrix_less_o3-fitted_5c_less_o3))
```

```{r, eval=TRUE, echo = T}
# NMF factorizes V = WH
# Store Basis matrix (W) and Coef Matrix (H)
saveRDS(basis_matrix_5c_less_o3, 'result_rfiles/nmf_norm_5c_less_o3_basis_diurnal.rds')
saveRDS(coef_matrix_5c_less_o3, 'result_rfiles/nmf_norm_5c_less_o3_coef_diurnal.rds')

# Merge basis matrix into hourly observations
basis_matrix_5c_less_o3 <- as_tibble(basis_matrix_5c_less_o3) %>%
  setNames(c('Factor1', 'Factor2', 'Factor3', 'Factor4', 'Factor5'))

normalized_hourly_data_diurnal_5c_less_o3_diurnal <- hourly_nona_diurnal[,c('day', 'time_utc')] %>%
  cbind(normalized_matrix_less_o3) %>%
  cbind(basis_matrix_5c_less_o3) %>%
  right_join(hourly_data_diurnal %>% select(-'day'), join_by(time_utc), suffix = c('_norm', ''))

# saveRDS(normalized_hourly_data_diurnal_5c_less_o3, 
#  'result_rfiles/normalized_hourly_data_diurnal_5c_less_o3_diurnal.rds')
normalized_hourly_data_5c_less_o3_diurnal <- readRDS('result_rfiles/normalized_hourly_data_diurnal_5c_less_o3_diurnal.rds')
```

- make daily dataset for VNF analysis 
- compute wind directions from plots

```{r}
# Also compute a daily dataset
normalized_daily_data_5c_less_o3_diurnal <- normalized_hourly_data_5c_less_o3_diurnal %>%
  group_by(day) %>%
  summarise(across(where(is.numeric) & !any_of('wdr_deg'), ~ mean(.x, na.rm = T)),
            wdr_deg = as.numeric(mean(circular(wdr_deg, units = "degrees"), na.rm = T))) %>%
  mutate(wdr_deg = if_else(wdr_deg < 0, wdr_deg+360, wdr_deg)) %>%
  mutate(wind_45_135 = wdr_deg >= 45 & wdr_deg < 135,
         wind_135_180 = wdr_deg >= 135 & wdr_deg < 180,
         wind_180_270 = wdr_deg >= 180 & wdr_deg < 270,
         wind_270_45 = wdr_deg >= 270 & wdr_deg < 45)

# saveRDS(normalized_daily_data_5c_less_o3_diurnal,
#  'result_rfiles/normalized_daily_data_5c_less_o3_diurnal.rds')

normalized_daily_data_5c_less_o3 <- 
  readRDS('result_rfiles/normalized_daily_data_5c_less_o3_diurnal.rds')
```

- 1) number of flares in 100km of trailer associated with NMF 
- 2) weighted count based on distance to trailer

```{r}
# Check if relationship between # flares and flare factor (4 & 5)
# Linear model
flare_factor <- lm(n_flare_100 ~ Factor1 + Factor2 + Factor3 + Factor4 + Factor5, 
                   data = normalized_daily_data_5c_less_o3)
summary(flare_factor)

flare_factor45 <- lm(n_flare_100 ~ Factor4 + Factor5, data = normalized_daily_data_5c_less_o3)
summary(flare_factor45)

flare_factor_weighted <- lm(weighted.count ~ Factor1 + Factor2 + Factor3 + Factor4 + Factor5, 
                            data = normalized_daily_data_5c_less_o3)
summary(flare_factor_weighted)

flare_factor_weighted45 <- lm(weighted.count ~ Factor4 + Factor5, 
                              data = normalized_daily_data_5c_less_o3)
summary(flare_factor_weighted45)
```

```{r}
# All factors + wind speed + wind direction + factor5:sw wind.
# Wind direction from 270 to 45 is left as reference group.
flare_factor_weighted_2 <- lm(weighted.count ~ Factor1 + Factor2 + Factor3 + 
                                Factor4 + Factor5 + wsp_ms + wind_45_135 + 
                                wind_135_180 + Factor5*wind_180_270, 
                              data = normalized_daily_data_5c_less_o3)
summary(flare_factor_weighted_2)

# Same as above but only factor 4 and 5
flare_factor_weighted_3 <- lm(weighted.count ~ Factor4 + Factor5 + wsp_ms + 
                                Factor5*wind_180_270, 
                              data = normalized_daily_data_5c_less_o3)
summary(flare_factor_weighted_3)

# Same as above but interaction between factor 4 and SW wind
flare_factor_weighted_3b <- lm(weighted.count ~ Factor4 + Factor5 + wsp_ms + 
                                 Factor4*wind_180_270, 
                               data = normalized_daily_data_5c_less_o3)
summary(flare_factor_weighted_3b)

# Same as above but with East wind
flare_factor_weighted_3c <- lm(weighted.count ~ Factor4 + Factor5 + wsp_ms + 
                                 Factor5*wind_45_135, 
                               data = normalized_daily_data_5c_less_o3)
summary(flare_factor_weighted_3c)

flare_factor_weighted_3d <- lm(weighted.count ~ Factor4 + Factor5 + wsp_ms + 
                                 Factor4*wind_45_135, 
                               data = normalized_daily_data_5c_less_o3)
summary(flare_factor_weighted_3d)

# Wind speed + factor 4 and interaction with East wind
flare_factor_weighted_4a <- lm(weighted.count ~ wsp_ms + Factor4*wind_45_135, 
                               data = normalized_daily_data_5c_less_o3)
summary(flare_factor_weighted_4a)

# Wind speed + factor 4 and interaction with SE wind
flare_factor_weighted_4b <- lm(weighted.count ~ wsp_ms + Factor4*wind_135_180, 
                               data = normalized_daily_data_5c_less_o3)
summary(flare_factor_weighted_4b)

# Wind speed + factor 4 and interaction with SW wind
flare_factor_weighted_4c <- lm(weighted.count ~ wsp_ms + Factor4*wind_180_270, 
                               data = normalized_daily_data_5c_less_o3)
summary(flare_factor_weighted_4c)

# Wind speed + factor 5 and interaction with East wind
flare_factor_weighted_5a <- lm(weighted.count ~ wsp_ms + Factor5*wind_45_135, 
                               data = normalized_daily_data_5c_less_o3)
summary(flare_factor_weighted_5a)

# Wind speed + factor 5 and interaction with SE wind
flare_factor_weighted_5b <- lm(weighted.count ~ wsp_ms + Factor5*wind_135_180, 
                               data = normalized_daily_data_5c_less_o3)
summary(flare_factor_weighted_5b)

# Wind speed + factor 5 and interaction with SW wind
flare_factor_weighted_5c <- lm(weighted.count ~ wsp_ms + Factor5*wind_180_270, 
                               data = normalized_daily_data_5c_less_o3)
summary(flare_factor_weighted_5c)

```

```{r}
# Check relationship between avg flare distance and flare factor (4 & 5)
# Linear model
flare_factor_dist <- lm(distToLovi ~ Factor4 + Factor5, data = normalized_daily_data_5c_less_o3)
summary(flare_factor_dist)
```
